% appendix; bring up result of proof and put details of
% transformation equations in a second section, along with
% the problem. Perhaps frame that later section as a
% discussion of two other ways of deriving 3.17 from 3.16
% that Einstein would seem to want to avoid. One which he
% does, and one which he doesn't. Also, begin this section
% by speaking of how Einstein calls the derivation a simple
% calculation, which in one sense is simple, but in another
% sense involved, and underlines the sense in which such a
% calculation is needed. Then frame the alternatives as ways
% in which the calculation wouldn't be needed, either
% because the transformation equations are simple; or
% because the compatibility proof assumes the conditional
% that Einstein is trying to proof.
\documentclass{article}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage[bookmarks]{hyperref}
\usepackage{enumerate}
\newtagform{brackets}{[}{]}
\usetagform{brackets}
\allowdisplaybreaks
\title{Einstein's Proof in \S3 of ``On the Electrodynamics of Moving Bodies''}
\author{Daniel Harrell}
\begin{document}
\maketitle

\textsc{``That Maxwell's electrodynamics}---the way in which
it is usually understood---when applied to moving bodies,
leads to asymmetries which do not appear to be inherent in
the phenomena is well known.''\footnote{All quotations are
	from the translation of the paper by Arthur
	I.~Miller in the Appendix of his book \emph{Albert
	Einstein's Special Theory of Relativity}, as are the
	numbers for all equations.} Thus begins Einstein's
1905 paper ``On the Electrodynamics of Moving
Bodies,'' the first presentation of the special theory of
relativity. After discussing an example of such 
asymmetry, Einstein writes this: 
\begin{quotation} 
	Examples of this sort, together with the
	unsuccessful attempts to discover any motion of the
	Earth relatively to the ``light medium,'' lead to
	the conjecture that to the phenomenon of absolute
	rest there correspond no properties of the
	phenomena, neither in mechanics, nor in
	electrodynamics, but rather that as has already been
	shown to quantities to the first order, for every
	reference system in which the laws of mechanics are
	valid, the laws of electrodynamics and optics are
	also valid.

	We will raise this conjecture (whose intent will
	from now on be referred to as the ``Principle of
	Relativity'') to a postulate, and moreover introduce
	another postulate, which is only apparently
	irreconcilable with the former: light is always
	propagated in empty space with a definite velocity
	$c$ which is independent of the state of motion of
	the emitting body. 
\end{quotation} 
The appearance of irreconcilability between these postulates
is evidently strong enough for Einstein, that he offers a
proof of their compatibility in \S3 of the
paper. Or so he claims---and my purpose in this essay will
be to determine whether the claim is true; and if so, in
what sense. Did Einstein indeed prove the postulates
compatible, and if so, how? 
\section{The Proof}
{\samepage
Einstein introduces his proof in \S3 as follows:
\begin{quote}
	We now have to prove that any ray of light, measured
	in the moving system, is propagated with the
	velocity $c$, if, as we have assumed, this is the
	case in the resting system; for we have not as yet
	furnished the proof that the principle of the
	constancy of the velocity of light is compatible
	with the principle of relativity.
\end{quote}
}
By introducing his proof this way, there seem to be at least
three questions that Einstein takes his paper to have 
already answered, along with a fourth we can raise at this
point to understand the necessity, or at least the
desirability, of such a proof: 
\begin{enumerate}
	\item What is the resting system?
	\item What is the moving system?
	\item Why would proving what Einstein says we now have to
		prove, about $c$'s being the velocity of any
		ray of light in the moving system if $c$ is
		its velocity in the resting system, prove in turn
		that the principle of this velocity's constancy 
		is compatible with the
		principle of relativity? 
	\item Why is the proof needed? That is, why might
		we think, until a proof shows us otherwise,
		that these principles were incompatible?
\end{enumerate}
It looks as if answers to all four questions can be
readily given as early as the beginning of \S2 of the paper,
given how Einstein formulates the definitions he offers
there of the principles in question (hereafter labelled
``P1'' and ``P2''):
\begin{quote}
\begin{enumerate}[P1.]
	\item The laws by which the states of physical systems
	undergo changes are independent of whether those
	changes of state are referred to one or the other of
	two coordinate systems moving relatively to each
	other in uniform translational
	motion.\label{P1}
	\item Any ray of light moves in the ``resting''
	coordinate system with the definite velocity $c$,
	which is independent of whether the ray was emitted
	by a resting or by a moving body.\label{P2}%Note later the
	% last part of this definition about the time
	%interval
\end{enumerate} 
\end{quote}
The resting system, then, is the coordinate
system in which any ray of light moves at $c$, independently
of any motion in the source of the ray, according to
P\ref{P1}. The
moving system, in turn, is any coordinate system in uniform
translational motion relative to the resting system; and
where, according to P\ref{P1}, either system can serve as reference
for any law of physics---in particular, the law stated in
P\ref{P2}. 

But then it would seem as if the principles were
incompatible. For suppose that the source of a ray of light
were moving at velocity $v$ in the direction of the ray.
This source would therefore move at $v$ in the resting
system (hereafter $K$), yet be at rest in a moving system
(hereafter $\kappa$): the system moving at $v$ in the
direction of the ray, relative to $K$. In both $K$ and
$\kappa$, according to P\ref{P1}, we should infer that this ray is
moving at $c$. But to infer it moves at $c$ in $\kappa$
would seem to violate P\ref{P2}. For it would then move at $c$
relative to its moving source, as if dependent on it.  And
according to P\ref{P2}, but now in violation of P\ref{P1}, it seems we
should say that the ray of light is moving at $c$ in $K$ but
$c-v$ in $\kappa$. Similarly, if the ray were moving
directly against the motion of its source, it seems we
should say that the ray is moving again at $c$ in $K$, but
now at $c+v$ in $\kappa$.  Proving the principles
compatible, then, would seem to mean showing that such
inferences from P\ref{P2} are wrong.  It would mean showing that if
the ray in question---and more generally, if \emph{any}
ray---is moving at $c$ in $K$ (according to
P\ref{P2}), it will indeed be moving at $c$ in $\kappa$
(in accordance with P\ref{P1}).  

And this is exactly what Einstein says we now have to prove
when he introduces his proof in \S3, quoted above.  On this
same level of description, moreover, the proof is
simple---and seemingly sound. Einstein supposes that light
is emitted in all directions from the origin of both $K$ and
$\kappa$. He further supposes, in accordance with P\ref{P2}, that
every ray is moving at $c$ in $K$, so forming
a spherical wave. Thus, he infers, for every point $(x,y,z)$
reached by the wave at time $t$ in $K$, the
distance $ct$ from that point back to origin will satisfy
the Pythagorean theorem in three dimensions: \begin{align}
	x^2+y^2+z^2=c^2 t^2.  \tag{\S3.16}\label{eq:3.16}
\end{align}

Now, if P\ref{P2} were incompatible with P\ref{P1}, we should not expect
this wave to be spherical in $\kappa$, but
lopsided, since we will expect the ray moving directly with
$\kappa$ to have a velocity of $c-v$, the ray
moving directly against this system a velocity of $c+v$, and
every other ray its own velocity between $c-v$ and $c+v$,
as determined by its direction. And in that case the wave would
\emph{not} satisfy the Pythagorean theorem \begin{align}
\xi^2 + \eta^2 + \zeta^2 = c^2\tau^2
\tag{\S3.17}\label{eq:3.17}\end{align}for every point
$(\xi,\eta,\zeta)$ reached at $\tau$ in $\kappa$,
according to its own spatio-temporal coordinates (and for
which Einstein uses the Greek letters given here). 
But it turns out that the wave \emph{does} satisfy this
theorem in $\kappa$. For when we actually transform
\ref{eq:3.16}, so that it refers to the coordinates of
$\kappa$ rather than those of $K$, the
result is \ref{eq:3.17}.
And this means that for every point
$(\xi,\eta,\zeta)$ reached by the wave at time $\tau$ in
$\kappa$, the distance from that point back to origin
is always $c\tau$. Or as Einstein characterizes the
seemingly successful result of the proof: \begin{quote} The
wave under consideration is therefore no less a
spherical wave with velocity of propagation $c$ when
considered in $\kappa$. Consequently our
two fundamental principles are compatible with each other.
\end{quote} 

\section{The Calculation of the Proof} 

Einstein says we obtain \ref{eq:3.17}
from \ref{eq:3.16} ``after a simple calculation.'' By
``simple,'' he seems to mean algebraic; for while the
calculation in question involves many steps, each step leads
to the next by algebraic manipulation alone.\footnote{The
calculation is laid out in Section
\ref{sec:Derivation}, p.~\pageref{sec:Derivation} of
this essay.} But that there is \emph{any}
calculation required is important, it seems, for
understanding the proof as indeed a proof. For
consider two ways that \ref{eq:3.17} might have
``followed from'' \ref{eq:3.16}, yet without
calculation. 

One way is if the equations relating coordinates between $K$
and $\kappa$ were as simple as this: \begin{align} \tau=t
	\quad \xi=x \quad \eta=y \quad
	\zeta=z\label{eq:simple} \end{align} For in that
case, \ref{eq:3.17} would follow from \ref{eq:3.16} as a
matter of simple substitution. But we know that $\xi\neq x$,
since the velocity $v$ of $\kappa$ displaces its
$\xi$-coordinates by at least the distance $vt$ relative to
$K$'s $x$-coordinates. Without any further complications,
this would mean that $x=\xi+vt$ and $x^2=\xi^2+2vt+v^2t^2$,
which when substituted for $x^2$ in \ref{eq:3.16}, would
evidently leave the irreducible quantity $2vt+v^2t^2$, to
preclude the derivation of \ref{eq:3.17} by any calculation.
But in a further complication, we also know that $\tau\neq
t$.  For as Einstein shows in \S2 of his paper, the simplest
definition of what it means for time-measuring ``clocks'' to
be synchronous in one system will preclude their being
synchronous in both systems at once. Moreover, this
preclusion follows from P\ref{P2}. For on this definition, any two
clocks $A$ and $B$ are synchronous if a ray of light is sent
from $A$ at $t_A$ to $B$ at $t_B$ and back to $A$ at $t'_A$,
and these times satisfy the equation: \begin{align}
	t_B-t_A=t'_A-t_B. \tag{\S1.1}\label{eq:1.1}
\end{align} But if $A$ and $B$ satisfy this equation while
at rest in $K$, they will no longer satisfy it when put in
motion with $\kappa$, by the same reasoning that makes
P\ref{P2}
seem incompatible with P\ref{P1}. For then the ray of light sent
between $A$ and $B$, in moving independently of their motion
by P\ref{P2}, will move in effect at $c-v$ from $A$ to $B$ and at
$c+v$ from $B$ back to $A$, with a corresponding difference
in the times to and back. Or as Einstein formulates it,
using $r_{AB}$ for the distance $AB$: \begin{align}
	t_B-t_A=\frac{r_{AB}}{c-v} \quad &\text{and} \quad
	t'_A-t_B=\frac{r_{AB}}{c+v}, \end{align} from which
it is clear that $t_B-t_A\neq t'_A-t_B$; or, for any clocks
$\alpha$ and $\beta$ that do satisfy \ref{eq:1.1} in
$\kappa$ (while $A$ and $B$ satisfy it in $K$), that
$\tau\neq t$.  So, since $\tau\neq t$ and $\xi\neq x$,
Einstein is forced in \S3 to derive a more complex set of
coordinate-relating equations than [\ref{eq:simple}]:
\begin{align} \tau&=\phi(v)\beta(t-vx/c^2)
	\tag{\S3.11}\label{eq:3.11}\\
	\xi&=\phi(v)\beta(x-vt)
	\tag{\S3.12}\label{eq:3.12}\\ \eta&=\phi(v)y
	\tag{\S3.13}\label{eq:3.13}\\ \zeta&=\phi(v)z
	\tag{\S3.14}\label{eq:3.14}, \end{align} where $c$
is the velocity of light in $K$, $v$ is the velocity of
$\kappa$ in relation to $K$, $\beta=1/(\sqrt{1-v^2/c^2)}$,
and $\phi(v)$ is a function that Einstein calls ``as yet
unknown,'' just before turning to his proof. The proof,
then, seems timed in \S3 for maximal effect; for it 
would seem improbable otherwise that
\ref{eq:3.17} could follow from \ref{eq:3.16} \emph{via}
\ref{eq:3.11}--\ref{eq:3.14} given
$\phi(v)$ alone, which is not simply a non-coordinate term
but an unknown one at that. And the ``known''
non-coordinate terms in \ref{eq:3.11} and \ref{eq:3.12},
which include  
$\beta$, $-vx/c^2$, and $-vt$, make the transformation at
first glance even
less probable.

Another way that \ref{eq:3.17} might have followed from
\ref{eq:3.16} without calculation is if we assumed
\ref{eq:3.16} by P\ref{P2}, and further assumed by P\ref{P1}
that if \ref{eq:3.16} then \ref{eq:3.17}.  \ref{eq:3.17}
would then follow from \ref{eq:3.16} as a simple matter of
logic: \begin{align*} &1. & x^2+y^2+z^2&=c^2 t^2
&\textrm{P2}\\ &2. & (x^2+y^2+z^2&=c^2 t^2) \rightarrow
(\xi^2 + \eta^2 + \zeta^2 = c^2\tau^2) &\textrm{P1}\\ &3. &
\therefore \quad \xi^2 + \eta^2 + \zeta^2 &= c^2\tau^2
&\textrm{1, 2\ Modus Ponens} \end{align*} This way of
deriving \ref{eq:3.17} from \ref{eq:3.16} (hereafter called
``logical''), however, begs the question of whether
P\ref{P1}, which is general, can indeed be applied to
P\ref{P2} to produce step (2) in the above proof.
Einstein's derivation, then, intends to prove rather than
assume (2), in order to settle rather than beg the question
of whether P\ref{P1} can be applied to P\ref{P2}.  And it is
in that sense that Einstein takes his derivation (hereafter
called ``material'') to prove their compatibility.

\section{The Problem} The difference between the material
and logical derivations of \ref{eq:3.17} from \ref{eq:3.16},
however, raises a question about Einstein's earlier
derivation of \ref{eq:3.11}--\ref{eq:3.14}.  For he derives
these latter equations by starting with the following
equations that he assumes to hold in $\kappa$: \begin{align}
	\tfrac{1}{2}(\tau_0+\tau_2)=\tau_1 \label{eq:tau}\\
\xi=c\tau \label{eq:xi}\\ \eta=c\tau \label{eq:eta}\\
\zeta=c\tau \label{eq:zeta} \end{align} The assumption
behind [\ref{eq:tau}] is that the clocks in $\kappa$ have
been synchronized according to \ref{eq:1.1}, while the
assumption behind [\ref{eq:xi}]--[\ref{eq:zeta}] is that
rays of light emitted from the origin of $\kappa$ along its
axes move at velocity $c$. But if this means that
\ref{eq:3.11}--\ref{eq:3.14} also assume so much about the
speed of light in $\kappa$, and it is by their application
alone that \ref{eq:3.17} is derived from \ref{eq:3.16}, then
it starts to look as if the material derivation, despite
requiring calculation, is no different in substance from the
logical derivation, where the application of
P\ref{P2} to P\ref{P1}, and thus their compatibility, is assumed rather
than proved. Moreover, Einstein includes a parenthetical
gloss on the source of [\ref{eq:xi}]--[\ref{eq:zeta}] that
suggests they come from the application of P\ref{P1} to
P\ref{P2}: ``\ldots [W]e can easily determine the quantities
$\xi$, $\eta$, $\zeta$ by expressing in equations that light
(as required by the principle of the constancy of the
velocity of light, in combination with the principle of
relativity) is also propagated with velocity $c$ when
measured in the moving system.'' 

His gloss, then, raises the question of
circularity in his supposed proof anew. For how has Einstein
indeed proved rather than merely assumed the compatibility
of P\ref{P2} and P\ref{P1}, if the derivation of \ref{eq:3.17} from
\ref{eq:3.16} that supposedly proves it depends on
transformation equations whose own derivation requires the
compatibility of P\ref{P2} and P\ref{P1} as an assumption from the start?
Understanding whether Einstein
indeed proved the compatibility of P\ref{P1} and P\ref{P2}, then, depends
on resolving this issue of circularity---the issue, that is,
of whether
\begin{align*}
	(x^2+y^2+z^2&=c^2 t^2) \rightarrow
(\xi^2 + \eta^2 + \zeta^2 = c^2\tau^2) 
\end{align*}
is indeed being proved by Einstein in his 
derivation of \ref{eq:3.17} from \ref{eq:3.16}, or rather
in some way being assumed.

\section{Solutions to the Problem} Unfortunately, Einstein
himself says nothing about this issue in the paper---a
surprising silence given how obvious the issue is. But this
suggests that Einstein's own understanding of his derivation
as a proof raises no such issue about circularity. So what
could this understanding be? I will now go through several
possible answers that all strike me as unlikely, essentially
because none of them resolves the issue of circularity in
such a way as to explain Einstein's silence about the
matter. But then, in the last part of this essay, I will 
propose a final answer that I think does resolve the issue
in this way, making the other resolutions even less likely.  

\subsection{An assumption of conjunction rather
compatibility} There are I will now go through what seem to
me And in that case Einstein has seemingly begged the very
question his proof was intended to settle. But how, in that
case, has Einstein not begged the question Indeed, if
\ref{eq:3.11}--\ref{eq:3.14} do assume that light moves at
$c$ in $\kappa$, then the only way it seems Einstein has
avoided assuming what he claims to prove, is if the
assumption that light moves at $c$ in $\kappa$ is not from
P\ref{P1}'s being applied to P\ref{P2}, but rather being made
independently. We could then interpret Einstein's
derivation, along with the calculation need to make it, as
showing the compatibility of P\ref{P1} and P\ref{P2} by combining, in
effect, the separate assumptions about the speed of light in
$K$ and $\kappa$ respectively. But it is hard to know what
could make the assumption that light moves at $c$ in
$\kappa$ truly independent from the assumption that light
moves at $c$ in $K$ rather than at best a silent extension
of it, since the latter assumption is also required for the
derivation. And in any case Einstein himself is not silent,
explaining the source of [\ref{eq:xi}]--[\ref{eq:zeta}]
before introducing them as if the assumption behind them
were indeed the application of P\ref{P1} to P\ref{P2}: ``\ldots~we can
easily determine the quantities $\xi$, $\eta$, $\zeta$ by
expressing in equations that light (as required by the
principle of the constancy of the velocity of light, in
combination with the principle of relativity) is also
propagated with velocity $c$ when measured in the moving
system.'' His gloss, then, raises the question of
circularity in his supposed proof anew. For how has Einstein
indeed proved rather than merely assumed the compatibility
of P\ref{P2} and P\ref{P1}, if the derivation of \ref{eq:3.17} from
\ref{eq:3.16} that supposedly proves it depends on
transformation equations whose own derivation requires the
compatibility of P\ref{P2} and P\ref{P1} as an assumption from the start?

that Einstein is indeed
assuming that these rays move at $c$ in $\kappa$ in deriving
, then how 
that cast his subsequent 
these assumptions, then,
make it harder to know why Einstein's derivation of 
$\ref{eq:3.17}$ from $\ref{eq:3.16}$ in a more suspicious
light. look suspiciously close to the question-begging
argumapproach of come suspiciously close to the
question-begging logical derivationapproach.

It is the
second way that \ref{eq:3.17} might have followed from
\ref{eq:3.16} without calculation, however, that portends a
problem with (Start revising from here.) from this one
Non-mathematically, The interpretation of this derivation is
also simple, given Einstein's suppositions for .
(\ref{eq:3.16}). means that light has traveled a distance
$ct$ at any point $(x,y,z)$ reached in a reference system
$K$ if (1) $K$ is taken to be at rest relative to another
reference system $\kappa$ taken to be in uniform translation
motion; (2) light is emitted from the origin of both sty
supposed at rest,results from a pair of suppositions: (1)
that we have emitted light in all directions The proof seems
to work as follows.  % Start here: The first sentence is
understandable enough. But what about the
% second one?  Start here, by raising a question about why
% this derivation, and more specifically, this derivation as
% Einstein does it through his transformation equations and
% simple calculation, counts as a proof of the compatibility
% of the postulates.

reached a reached by the wave ere 

( For anyne way of
understanding the first part of the paper proper, from \S1 through little more
than halfEinstein offers his ``proof'' that the two postulates are compatible,
he does it by deriving this equation,


from this one,
Einstein calls the calculation behind this derivation ``simple''. But there
would seem a still simpler proof of compatibility, involving no such
calculation in the derivation. 

First, we assume that light moves at $c$ in resting system $K$, as
required by the principle of the constancy of the velocity of light. Thus,
a ray of light emitted along the axes of $K$ will satisfy the following
equations: 
\begin{align*}
	x=ct\\
	y=ct\\
	z=ct
\end{align*}
Moreover, any other ray of light emitted from the origin of $K$ can be 
expressed by the ordered triple of $K$ co-ordinates $(x,y,z)$, and
understood to satisfy the distance formula $x^2+y^2+z^2=c^2t^2$.

Next, we assume that if light moves at $c$ in resting system $K$, then it
moves at $c$ in moving system $\kappa$, as 
required by the principle of relativity. Thus, if a ray of light emitted along
the axes of $K$ satisfied the previous equations, then a ray of light emitted along the
axes of $\kappa$ will satisfy the following equations:
\begin{align*} \xi=c\tau\\
\eta=c\tau\\ \zeta=c\tau \end{align*}

Moreover, any other ray of light emitted
from the origin of $\kappa$ can be expressed by the ordered triple of $\kappa$
coordinates $(\xi,\eta,\zeta)$, and understood to satisfy the distance formula
$\xi^2+\eta^2+\zeta^2=c^2\tau^2$.

Our two assumptions, then, are (1) that this equation holds:
\begin{align*}
x^2+y^2+z^2=c^2 t^2;
\end{align*}
and (2) that if that equation holds, then this equation holds:
\begin{align*}
\xi^2 + \eta^2 + \zeta^2 = c^2\tau^2;
\end{align*}
from which it follows, taking both assumptions together, that the latter
equation holds:
\begin{align*}
\xi^2 + \eta^2 + \zeta^2 = c^2\tau^2.
\end{align*}

% Go on to ask why Einstein bothers with the calculation. Use quote to show
% he is taking the compatibility proof to be of the second assumption, that is,
% to assume the proposition required by one postulate, and prove the proposition
% required by the other postulate. But why do that? Leave that question for
% later, go on to discuss circularity.

But why, then, does Einstein bother with a derivation? Presumably 
because he takes a compatibility proof of the postulates to require more than
the mere assumption of both postulates at once. And he seems to say as much
when he introduces the proof:
\begin{quote}
	We now have to prove that any ray of light, measured in the moving
	system, is propagated with velocity $c$, if, as we have assumed,
	this is the case in the resting system; for we have not yet
	furnished the proof that the principle of the constancy of the
	velocity of light is compatible with the principle of relativity.
\end{quote}
It looks as if Einstein takes a compatibility proof in this case to mean
that we assume only that light moves at $c$ in resting system $K$, as required by
the principle of the constancy of the velocity of light, but \emph{prove} on this
assumption that light moves at $c$ in moving system $\kappa$, as required by the
principle of relativity. ssume the principle of the constancy of the velocity of light in
the resting system, prove
	Or to put the point in the
terms that he himself uses to introduce the proof: he takes a compatibility
proof in this case to require that we assume only that light move at $c$ in
$K$, but \emph{prove} on that assumption that light move at $c$ in $\kappa$.

Looks as if we can assume both without contradiction, so the two principles
are compatible. Why doesn't Einstein do it this way? Without the derivation.

Note as well that the derivation seems to be if light moves at c in K,
it will moves at c in kappa. We don't assume both at once, but just that
light moves at c in K, then "find out" through the derivation that it
travels at c in kappa.
Answer, there is something in the first principle left out of the proof, that
makes the principles seem to be incompatible.

Circularity would seem to be that we assume light travels at c in kappa. But
Einstein assumes this in his actual approach.

Put the point this way: Einstein makes no assumption that the principle
of relativity holds for the transformation equations insofar as they
are transformation equations. Instead the assumption is: if light
travels at velocity c in kappa as measured in kappa, then it will not travel
at c as measured in K. Or the other way around, if light travels at
velocity c in kappa as measured in K, it will not travel at c in kappa
as measured in K. 

\newpage
\appendix
\section{Derivation}\label{sec:Derivation}
We begin with \ref{eq:3.16}: 
\begin{align}  x^2+y^2+z^2=c^2 t^2.
\tag{\ref{eq:3.16}} \end{align}
Then reverting back to equations 
\ref{eq:3.11}--\ref{eq:3.14},
 we solve for the $K$
coordinates as follows:
\setcounter{equation}{0}
\begin{align}
	\tau&=\phi(v)\beta\left(t-\frac{vx}{c^2}\right)
	\tag{\ref{eq:3.11}}\\
 \tau&=\phi(v)\beta t - \phi(v)\beta\frac{vx}{c^2}\\
\phi(v)\beta\frac{vx}{c^2}+\tau&=\phi(v)\beta t\\
	\frac{vx}{c^2}+\frac{\tau}{\phi(v)\beta}&=t\label{eq:t}\\[1em]
	\xi&=\phi(v)\beta(x-vt) \tag{\ref{eq:3.12}}\\
\xi&=\phi(v)\beta x -\phi(v)\beta vt\\
\phi(v)\beta vt+\xi&=\phi(v)\beta x\\
	vt+\frac{\xi}{\phi(v)\beta}&=x\label{eq:x}\\[1em]
	\eta&=\phi(v)y \tag{\ref{eq:3.13}}\\
	\frac{\eta}{\phi(v)}&=y\label{eq:finaly}\\[1em]
	\zeta&=\phi(v)z \tag{\ref{eq:3.14}}\\
\frac{\zeta}{\phi(v)}&=z\label{eq:finalz}
\end{align}
To eliminate the embedded $K$ coordinates that remain in
equations [\ref{eq:t}] and [\ref{eq:x}], we can substitute
[\ref{eq:x}] for $x$ in [\ref{eq:t}], and [\ref{eq:t}] for $t$
in [\ref{eq:x}]. Solving again for $t$:
\begin{align}
	\frac{vx}{c^2}+\frac{\tau}{\phi(v)\beta}&=t\tag{\ref{eq:t}}\\
  \frac{v\left(vt+\frac{\xi}{\phi(v)\beta}\right)}{c^2}+\frac{\tau}{\phi(v)\beta}&=t\\
  \frac{v^2t+\frac{v\xi}{\phi(v)\beta}}{c^2}+\frac{\tau}{\phi(v)\beta}&=t\\
  \frac{v^2t}{c^2}+\frac{v\xi}{\phi(v)\beta c^2}+\frac{\tau}{\phi(v)\beta}&=t\\
  \frac{v\xi}{\phi(v)\beta c^2}+\frac{\tau}{\phi(v)\beta}&=t-\frac{v^2t}{c^2}\\
  \frac{v\xi}{\phi(v)\beta c^2}+\frac{\tau}{\phi(v)\beta}&=t\left(1-\frac{v^2}{c^2}\right)\\
  \frac{1}{1-\frac{v^2}{c^2}}\left(\frac{v\xi}{\phi(v)\beta c^2}+\frac{\tau}{\phi(v)\beta}\right)&=t\\
  \beta^2\left(\frac{v\xi}{\phi(v)\beta c^2}+\frac{\tau}{\phi(v)\beta}\right)&=t\\
  \frac{\beta v\xi}{\phi(v)c^2}+\frac{\beta\tau}{\phi(v)}&=t\\
  \intertext{or} \frac{\frac{\beta
      v\xi}{c^2}+\beta\tau}{\phi(v)}&=t\\
  \frac{\beta(\frac{v\xi}{c^2}+\tau)}{\phi(v)}&=t\label{eq:finalt}.
\end{align}
And solving again for $x$:
\begin{align}
	vt+\frac{\xi}{\phi(v)\beta}&=x\tag{\ref{eq:x}}\\
v\left(\frac{vx}{c^2}+\frac{\tau}{\phi(v)\beta}\right)+\frac{\xi}{\phi(v)\beta}&=x\\
\frac{v^2x}{c^2}+\frac{v\tau}{\phi(v)\beta}+\frac{\xi}{\phi(v)\beta}&=x\\
\frac{v\tau}{\phi(v)\beta}+\frac{\xi}{\phi(v)\beta}&=x-\frac{v^2x}{c^2}\\
\frac{v\tau}{\phi(v)\beta}+\frac{\xi}{\phi(v)\beta}&=x\left(1-\frac{v^2}{c^2}\right)\\
\frac{1}{1-\frac{v^2}{c^2}}\left(\frac{v\tau}{\phi(v)\beta}+\frac{\xi}{\phi(v)\beta}\right)&=x\\
\beta^2\left(\frac{v\tau}{\phi(v)\beta}+\frac{\xi}{\phi(v)\beta}\right)&=x\\
\frac{\beta v\tau + \beta \xi}{\phi(v)}&=x\\
\frac{\beta(v\tau + \xi)}{\phi(v)}&=x\label{eq:finalx}
\end{align}

Now we plug [\ref{eq:finalx}], [\ref{eq:finalt}],
[\ref{eq:finalz}], and [\ref{eq:finaly}] back into
\ref{eq:3.16}:
\begin{align}  x^2+y^2+z^2&=c^2 t^2
\tag{\ref{eq:3.16}}\\
  \left(\frac{\beta(v\tau +
      \xi)}{\phi(v)}\right)^2+\left(\frac{\eta}{\phi(v)}\right)^2+\left(\frac{\zeta}{\phi(v)}\right)^2&=c^2\left(\frac{\beta(\frac{v\xi}{c^2}+\tau)}{\phi(v)}\right)^2\\
  \frac{\beta^2(v\tau +
    \xi)^2}{\phi^2(v)}+\frac{\eta^2}{\phi^2(v)}+\frac{\zeta^2}{\phi^2(v)}&=c^2\left(\frac{\beta^2(\frac{v\xi}{c^2}+\tau)^2}{\phi^2(v)}\right)\\
  \left(\frac{\phi^2(v)}{\beta^2(\frac{v\xi}{c^2}+\tau)^2}\right)\left(\frac{\beta^2(v\tau + \xi)^2+\eta^2+\zeta^2}{\phi^2(v)}\right)&=c^2\\
  \left(\frac{\phi^2(v)}{\beta^2(\frac{v\xi}{c^2}+\tau)^2}\right)\left(\frac{\beta^2\left[(v\tau + \xi)^2+\frac{\eta^2}{\beta^2}+\frac{\zeta^2}{\beta^2}\right]}{\phi^2(v)}\right)&=c^2\\
\frac{(v\tau +
    \xi)^2+\frac{\eta^2}{\beta^2}+\frac{\zeta^2}{\beta^2}}{(\frac{v\xi}{c^2}+\tau)^2}&=c^2\\
  (v\tau +
  \xi)^2+\frac{\eta^2}{\beta^2}+\frac{\zeta^2}{\beta^2}&=c^2\left(\frac{v\xi}{c^2}+\tau\right)^2\\
  (v\tau +
  \xi)^2+\frac{\eta^2}{\beta^2}+\frac{\zeta^2}{\beta^2}&=c^2\left(\frac{v^2\xi^2}{c^4}+\frac{2\tau
    v \xi}{c^2} + \tau^2 \right)\\
  (v\tau +
  \xi)^2+\frac{\eta^2}{\beta^2}+\frac{\zeta^2}{\beta^2}&=\frac{v^2\xi^2}{c^2}+2\tau
    v \xi + c^2\tau^2\\
v^2\tau^2 +
  2\tau v \xi + \xi^2+\frac{\eta^2}{\beta^2}+\frac{\zeta^2}{\beta^2}&=\frac{v^2\xi^2}{c^2}+2\tau
    v \xi + c^2\tau^2\\
v^2\tau^2 +
  2\tau v \xi +
  \xi^2-\frac{v^2\xi^2}{c^2}-2\tau
    v \xi + \frac{\eta^2}{\beta^2}+\frac{\zeta^2}{\beta^2}&=c^2\tau^2\\
v^2\tau^2 +
 \xi^2-\frac{v^2\xi^2}{c^2}+ \frac{\eta^2}{\beta^2}+\frac{\zeta^2}{\beta^2}&=c^2\tau^2\\ 
\xi^2-\frac{v^2\xi^2}{c^2}+ \frac{\eta^2}{\beta^2}+\frac{\zeta^2}{\beta^2}&=c^2\tau^2-v^2\tau^2\\
\xi^2\left(1-\frac{v^2}{c^2}\right)+\frac{\eta^2}{\beta^2}+\frac{\zeta^2}{\beta^2}&=c^2\tau^2-v^2\tau^2\\
\xi^2\left(1-\frac{v^2}{c^2}\right)+\frac{\eta^2}{\beta^2}+\frac{\zeta^2}{\beta^2}&=c^2\tau^2\left(1-\frac{v^2}{c^2}\right)\\
\xi^2\left(\frac{1}{\beta^2}\right)+\frac{\eta^2}{\beta^2}+\frac{\zeta^2}{\beta^2}&=c^2\tau^2\left(\frac{1}{\beta^2}\right)\\
\frac{\xi^2}{\beta^2}+\frac{\eta^2}{\beta^2}+\frac{\zeta^2}{\beta^2}&=\frac{c^2\tau^2}{\beta^2}\\
\xi^2+\eta^2+\zeta^2&=c^2\tau^2. \tag{\ref{eq:3.17}}
\end{align}
\end{document}

